---
title: "week1_hw"
author: "Eva Newby"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(sf)
library(here)
library(terra)

# read in major points on the island
ei_points <- sf::read_sf(here::here("data", "easter-island", "ei_points.gpkg"))
# subset points to volcanoes
volcanoes <- subset(ei_points, type == "volcano")

# read in island elevation
ei_elev <- terra::rast(here::here("data", "easter-island", "ei_elev.tif"))

# read in island border
ei_borders <- sf::read_sf(here::here("data", "easter-island", "ei_border.gpkg"))

# read in island road network
ei_roads <- sf::read_sf(here::here("data", "easter-island", "ei_roads.gpkg"))

# read in bathymetry
ei_bathymetry <- terra::rast(here::here("data", "easter-island", "ei_bathy.tif"))

# read in seamounts
ei_seamounts <- sf::read_sf(here::here("data", "easter-island", "ei_seamounts.gpkg"))
```

Description
Create a map of Easter Island, including the following:

- the island’s border and road network
- elevation of the island (indicated in meters above sea level)
- location of the island’s volcanoes (indicating their height in meters)
- bathymetry surrounding the island (indicated in meters below sea level)
-surrounding seamounts (indicating their height in meters)
```{r}
  tm_shape(ei_bathymetry)+ #bathymetry
  tm_raster(palette = c("lightblue", "darkblue"), # need help with my color palette here, looks odd
            title = 'Bathymetry (m below sea level)')+
tm_shape(ei_elev)+
  tm_raster(palette = c("#0B8A26", "#D4EBD9"),
            title = 'Elevation (m amsl)')+  # map of elevation
  tm_shape(ei_borders)+
  tm_borders()+  # map of Easter island
  tm_shape(ei_roads)+
  tm_lines()+ # roads
tm_shape(volcanoes)+
  tm_dots(col = '#EF476F', size = .1)+  # volcanoes (need height in m)
  tm_shape(ei_seamounts)+
  tm_dots(col = 'orange', size = .1)+
  tm_scale_bar(position = c('left', 'bottom'))+
  tm_compass() +
  tm_layout(bg.color = 'grey95',
            main.title = 'Elevation on Easter Island',
            frame = FALSE,
            main.title.size = 1)
```

